% =====================================================================
% PAPER III — Eisenstein–Picard by verified finite elements
% Level-1 gap for PSL(2,Z[omega]) and congruence rungs N=3,7
% ---------------------------------------------------------------------
% Sources: eisenstein_numbers/{PROOF,PAIRING_MATRICES,T0_AREA,GEOMETRY,
%          CERT_OMEGA_P}.md; cert_omega.py, cert_omega_p.py
% Parallel method paper: papers/paper1_hammer.tex (Picard level 1)
% =====================================================================
\documentclass[11pt]{amsart}
\usepackage{amsmath,amssymb,amsthm,booktabs}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{corollary}[theorem]{Corollary}
\newcommand{\HH}{\mathbb{H}^3}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\PSL}{\mathrm{PSL}}
\newcommand{\vol}{\mathrm{vol}}
\newcommand{\p}{\mathfrak{p}}

\title[Verified exclusion for the Eisenstein--Picard orbifold]{Exceptional
eigenvalues of Bianchi orbifolds by verified finite elements.\\
III: The Eisenstein--Picard orbifold and congruence covers}
\author{}
\date{Draft, 2026-07-12}

\begin{document}
\begin{abstract}
We apply the verified finite-element exclusion method of Paper~I to the
Eisenstein--Picard group $\Gamma=\PSL(2,\ZZ[\omega])$, $\omega=e^{2\pi i/3}$.
Using the Elstrodt--Grunewald--Mennicke fundamental domain, Crouzeix--Raviart
elements on a truncated core, ball-arithmetic enclosures, and Rump's
positive-definiteness certificate, we prove that
$L^2(\Gamma\backslash\HH)$ has no Laplace eigenvalue in $(0,1)$.  The same
architecture, with multi-copy reference cells and two-cusp weights, yields
the first congruence rungs: $\Gamma_0(\p)\backslash\HH$ for
$N(\p)\in\{3,7\}$ likewise has no eigenvalue in $(0,1)$.  The method is
independent of the Selberg trace formula.  All certificates are
deterministic and reproducible.
\end{abstract}
\maketitle

% ---------------------------------------------------------------------
\section{Introduction}

Bianchi groups $\PSL(2,\mathcal{O}_K)$ for imaginary quadratic
$\mathcal{O}_K$ are the natural $\mathrm{GL}(2)$ arithmetic subgroups
acting on hyperbolic $3$-space.  Papers~I and~II of this series treated
the Gaussian case $K=\mathbb{Q}(i)$: Paper~I developed the verified
finite-element exclusion method and proved $\lambda_1\ge 1$ for the Picard
orbifold; Paper~II climbed the congruence ladder
$\Gamma_0(\p)$ for Gaussian primes of norms $5$, $9$, and $13$.  The
present paper carries the same architecture to the Eisenstein integers
$\mathcal{O}=\ZZ[\omega]$, $\omega=e^{2\pi i/3}$, the ring of integers of
$K=\mathbb{Q}(\sqrt{-3})$.

The Eisenstein--Picard group $\Gamma=\PSL(2,\ZZ[\omega])$ is, like the
Picard group, a class-number-one Bianchi group with a single cusp.  Its
fundamental domain is smaller (covolume $\approx 0.169$ versus
$\approx 0.305$), the cusp stabilizer includes a rotation of order $3$
rather than $2$, and the planar cross-section is the EGM domain $P_3$ of
area $\sqrt{3}/6$ rather than the Humbert rectangle of area $1/2$.  These
are geometric differences; the analytic criterion, the guaranteed FEM
lower bound (Theorem~G1), and the Rump certificate transfer verbatim once
the fundamental domain, the cross-section area, and the pairing dictionary
are correctly installed.

\begin{theorem}[Level 1]\label{thm:A}
The Laplace--Beltrami operator on
$L^2(\PSL(2,\ZZ[\omega])\backslash\HH)$ has no eigenvalue in $(0,1)$.
Combined with the spectral decomposition of cofinite Kleinian groups
\cite[Thm.~3.8.1]{Friedman}, $\lambda_1\ge 1$.
\end{theorem}

\begin{theorem}[Congruence rung $N=3$]\label{thm:B}
Let $\p=(1-\omega)\subset\ZZ[\omega]$ (so $N(\p)=3$) and
\[
  \Gamma_0(\p)
  =\Bigl\{\begin{pmatrix}a&b\\c&d\end{pmatrix}\in\PSL(2,\ZZ[\omega])
  :\;c\in\p\Bigr\}.
\]
Then $[\Gamma:\Gamma_0(\p)]=N(\p)+1=4$, the quotient
$\Gamma_0(\p)\backslash\HH$ has exactly two cusps ($\infty$ and $0$), and
the Laplacian on $L^2(\Gamma_0(\p)\backslash\HH)$ has no eigenvalue in
$(0,1)$.
\end{theorem}

\begin{theorem}[Congruence rung $N=7$]\label{thm:Btwo}
Let $\p\subset\ZZ[\omega]$ be a prime ideal of norm $N(\p)=7$ (so
$\mathcal{O}/\p\cong\mathbb{F}_7$).  Then $[\Gamma:\Gamma_0(\p)]=8$,
the quotient has two cusps with $|T_\infty|=\sqrt{3}/6$ and
$|T_0|=7\cdot|T_\infty|$, and the Laplacian on
$L^2(\Gamma_0(\p)\backslash\HH)$ has no eigenvalue in $(0,1)$.
\end{theorem}

These theorems are independent of the Selberg trace formula.  The companion
repository also certifies a trace-formula bound $B(\mathbb{Q}(\omega))<1$
at level one, by a completely disjoint method sharing only the fundamental
domain; Theorem~\ref{thm:A} is therefore an independent cross-check of the
level-one spectral gap.  Theorems~\ref{thm:B}--\ref{thm:Btwo} have no
trace-formula counterpart in the repository: as in Paper~II, the
$B$-constant degrades with volume and the FEM path is the one that reaches
congruence level.

\begin{remark}[What is not claimed]
No statement is made for $N(\p)=13$ over $\ZZ[\omega]$ beyond
combinatorics (residue arithmetic and gluing tables exist; interval
certificates do not).  No re-proof of the Gaussian ladder
(Paper~II, Theorems for $N\p=5,9,13$) is attempted.  The certificates
below do not use the Selberg trace formula, any test function, the
scattering matrix, or the torsion inventory.
\end{remark}

Reproduce (from \texttt{eisenstein\_numbers/}):
\begin{verbatim}
python -u cert_omega.py 6 3         # Theorem A
python -u cert_omega_p.py 3 6 3     # Theorem B (N=3)
python -u cert_omega_p.py 7 6 3     # Theorem B2 (N=7)
\end{verbatim}

% ---------------------------------------------------------------------
\section{Geometry}\label{sec:geom}

\subsection{Group and fundamental domain}

Let $\omega=e^{2\pi i/3}=-\tfrac12+i\sqrt{3}/2$, so
$\mathcal{O}=\ZZ[\omega]=\ZZ\oplus\ZZ\omega$ with covolume
$\mathrm{covol}_{\RR^2}(\mathcal{O})=\sqrt{3}/2$ and unit group
$\mathcal{O}^\times=\{\pm 1,\pm\omega,\pm\omega^2\}$ of order $6$.  In
$\PSL$, $-I$ is trivial, so the image of $\mathcal{O}^\times$ in the cusp
stabilizer acts on $\CC$ by rotations of order $3$.  Write
$U=\langle z\mapsto\omega z\rangle\subset\mathrm{Isom}(\CC)$, $|U|=3$.
The parabolic stabilizer is the semidirect product
$\Gamma_\infty\cong\mathcal{O}\rtimes U$.

The fundamental polyhedron is that of Elstrodt--Grunewald--Mennicke
\cite[§7.2--7.3]{EGM98} (restated as \cite[Thm.~2.3]{DP20}):
\[
  F_3=\bigl\{(z,t)\in B_3:\;z\in P_3\bigr\},
\]
where $B_3$ is the Ford region (for $d=3$, the only exterior sphere cutting
the prism $P_3\times\RR_{>0}$ is the unit hemisphere) and $P_3$ is the
planar domain
\begin{align*}
  P_3&=T_{\mathrm{up}}\cup T_{\mathrm{low}},\\
  T_{\mathrm{up}}
  &=\bigl\{x+iy:\;0\le x,\; x/\sqrt{3}\le y\le(1-x)/\sqrt{3}\bigr\}
  \quad(x\le\tfrac12),\\
  T_{\mathrm{low}}
  &=\bigl\{x+iy:\;0\le x\le\tfrac12,\; -x/\sqrt{3}\le y\le x/\sqrt{3}\bigr\}.
\end{align*}
Euclidean area: $\mathrm{area}(P_3)=1/(2\sqrt{3})=\sqrt{3}/6$.  Since
$\max_{P_3}|z|=1/\sqrt{3}<1$,
\[
  F_3=\bigl\{(z,t):\;z\in P_3,\; t\ge\sqrt{1-|z|^2}\bigr\}.
\]
The orbifold cusp cross-section has area
\[
  |T|=\frac{\mathrm{covol}(\mathcal{O})}{|U|}=\frac{\sqrt{3}/2}{3}=\frac{\sqrt{3}}{6}.
\]

\subsection{Truncated core and volume}

Fix truncation height $Y>1$ (default $Y=1.25$).  The compact core is
\[
  K_Y=F_3\cap\{t\le Y\}
  =\bigl\{(z,y):\;z\in P_3,\; y_f(z)\le y\le Y\bigr\},
  \qquad y_f=\sqrt{1-|z|^2}.
\]
The Humbert volume formula gives
\[
  \vol(F)=\frac{|D|^{3/2}\zeta_K(2)}{4\pi^2}
  =\frac{3\sqrt{3}\,\zeta_K(2)}{4\pi^2}
  \approx 0.169156934402
\]
($D=-3$, $\zeta_K(2)=\zeta(2)L(2,\chi_{-3})$).  The cusp tail has product
structure $F\cap\{t>Y\}\cong P_3\times(Y,\infty)$ with hyperbolic measure
$t^{-3}\,dt\,dx\,dy$, so
\[
  \vol(F\cap\{t>Y\})=\frac{|T|}{2Y^2},\qquad
  \vol(K_Y)=\vol(F)-\frac{|T|}{2Y^2}.
\]
At $Y=1.25$: tail $\approx 0.092376$, $\vol(K_Y)\approx 0.076781$.

\subsection{Pairing matrices}

Generators in $\mathrm{SL}(2,\ZZ[\omega])$, descending to $\Gamma$:
\begin{align*}
  T_1&=\begin{pmatrix}1&1\\0&1\end{pmatrix},&
  T_\omega&=\begin{pmatrix}1&\omega\\0&1\end{pmatrix},\\
  U&=\begin{pmatrix}\omega^2&0\\0&\omega\end{pmatrix},&
  S&=\begin{pmatrix}0&-1\\1&0\end{pmatrix}.
\end{align*}
Orders in $\PSL$: $T_1,T_\omega$ parabolic; $U$ of order $3$; $S$ of order
$2$.  Inverses: $T_1^{-1}=\bigl(\begin{smallmatrix}1&-1\\0&1\end{smallmatrix}\bigr)$,
$T_\omega^{-1}=\bigl(\begin{smallmatrix}1&-\omega\\0&1\end{smallmatrix}\bigr)$,
$U^{-1}=\bigl(\begin{smallmatrix}\omega&0\\0&\omega^2\end{smallmatrix}\bigr)$,
$S^{-1}=S$.

Reference face maps on the computational cell over $P_3$:
\begin{center}
\begin{tabular}{@{}lll@{}}
\toprule
Generator $\delta$ & Edges / faces paired & Reference map \\
\midrule
$T_1$ & RIGHT ($x=\tfrac12$) $\leftrightarrow$ LEFT ($x=0$)
  & height-matched opposite sides \\
$T_\omega$ & LOW $\leftrightarrow$ UP (slanted sides)
  & height-matched opposite slants \\
$U$ & vertical $\leftrightarrow$ vertical
  & $\varphi_U(z,t)=(\omega z,t)$ \\
$S$ & FLOOR $\leftrightarrow$ FLOOR
  & full $\HH$ action (sphere inversion) \\
--- & TOP ($t=Y$) & free (cusp ODE / Lemma~D0) \\
\bottomrule
\end{tabular}
\end{center}

Ideal-vertex cycles of the inverses (polyhedron section; machine-checked
against a fixed $\Gamma_\infty$-section):
\begin{align*}
  T_1^{-1}&:\;(\infty)(0\ 1)(\omega\ \omega^2),\\
  T_\omega^{-1}&:\;(\infty)(0\ \omega)(1\ \omega^2),\\
  U^{-1}&:\;(\infty)(0)(1\ \omega\ \omega^2)
  \quad\text{(equals forward $U$ as a cycle; same $\langle U\rangle$)},\\
  S^{-1}&:\;(\infty\ 0)(1)(\omega\ \omega^2).
\end{align*}

\begin{lemma}[Pairing matrices]\label{lem:pair}
\begin{enumerate}
\item The four matrices above lie in $\mathrm{SL}(2,\mathcal{O})$ and
  represent elements of $\Gamma$.
\item For every floor face $f$ in the certified mesh dictionary, $S(f)$
  lands on the dictionary partner (within mesh tolerance); $S^2=\mathrm{id}$.
\item For every vertical dictionary pair labelled $U$, the image of the
  source centre under $z\mapsto\omega z$ matches the partner centre up to
  mesh tolerance.
\item $T_1$ and $T_\omega$, together with $U$, generate the cusp
  stabilizer $\Gamma_\infty$; the $P_3$ edge dictionary is the
  combinatorial section used for residue gluing.
\item Residue right action by $\delta^{-1}$ on $\mathbb{P}^1(\mathcal{O}/\p)$
  is bijective, with $S^2=U^3=\mathrm{id}$ (checked for
  $N\in\{3,7,13\}$ at the combinatorial level).
\end{enumerate}
Parts (1)--(3) and (5) are machine-checked; part (4) records the EGM/Swan
section convention.
\end{lemma}

\begin{remark}[Neumann versus non-Neumann]
The certified level-one FE space uses Neumann free $H^1(K)$ (relaxation:
larger space).  As in Paper~I, positivity on free $H^1(K)$ implies the
same on the subspace of functions that descend to the quotient.  An
optional module imposes the pairing dictionary on CR face dofs (raw
$1440\to$ paired $1342$ dofs at the certified mesh); positivity there is
implied by the Neumann certificate and is checked in floating point for
both $\kappa_{\mathrm{sc}}$ (Lemma~I1 of Paper~I) and the sharper CZZ
constant, with $\min c_e/d_e\approx 1.53$ (I1) and $\approx 2.09$ (CZZ).
This optional run is a geometric check, not a second interval
certificate.
\end{remark}

\subsection{Shimizu for $\ZZ[\omega]$}

The precisely-invariant-horoball argument is identical to Paper~I once one
notes that the minimal nonzero ideal norm in $\mathcal{O}$ is $1$: for
$\gamma\in\Gamma$ with bottom-left entry $c\neq 0$,
$N(c)\ge 1$, so $y(\gamma P)\le 1/(|c|^2 y)\le 1/y$.  Hence $\{y>Y\}$ with
$Y\ge 1$ meets its $\Gamma$-translates only through $\Gamma_\infty$.

% ---------------------------------------------------------------------
\section{Analytic criterion}\label{sec:crit}

\subsection{Level one}

The Lax--Phillips analysis of Paper~I, Sections~2--3, applies with the
single change $|T|=\sqrt{3}/6$ in place of $|T|=1/2$.  For
$v\in H^1(K_Y)$ and $s\in(0,1)$, $\lambda=1-s^2$:
\begin{align*}
  \mathcal{A}_s(v)
  &=Q_{K_Y}(v)-\lambda M_{K_Y}(v)-\beta(s)\,t(v)^2,\\
  \mathcal{L}_s(v)
  &=a(v)+\kappa_c(s)\,t(v),
\end{align*}
where $t(v)=\int_{P_3}v(\cdot,Y)\,dx\,dy$ (Euclidean),
$a(v)=\int_{K_Y}v\,y^{-3}\,dV$, and
\begin{equation}\label{eq:beta1}
  \beta(s)=\frac{1-s}{|T|Y^2},\qquad
  \kappa_c(s)=\frac{1}{(1+s)Y^2}.
\end{equation}
(The factor $1/|T|$ in $\beta$, rather than Paper~I's factor $2$ coming
from $|T|=1/2$ and a double-cover convention, is the natural single-cusp
normalization with cross-section area $|T|$.)  Nonzero Fourier modes in
the cusp are nonnegative: the dual lattice $\Lambda^*$ has minimal length
$2/\sqrt{3}$, so $4\pi^2 Y^2/|\mu|_{\min}^2\gg 1$ at $Y=1.25$.  The
zero-mode ODE and orthogonality constraint are as in Paper~I.  The
exclusion criterion and pencil form (Paper~I, Theorem on the criterion
and Lemma~P) transfer verbatim.

\begin{theorem}[Level-one criterion]\label{thm:critA}
If for every $s\in(0,1)$ the form $\mathcal{A}_s$ is positive on
$\{\mathcal{L}_s=0\}\subset H^1(K_Y)$, then
$\PSL(2,\ZZ[\omega])\backslash\HH$ has no eigenvalue in $(0,1)$.
\end{theorem}

Testing over all of $H^1(K_Y)$ is a relaxation (no side/floor
identifications).  Continuous spectrum of cofinite Kleinian groups starts
at $1$ \cite[Thm.~3.8.1]{Friedman}; atoms in $(0,1)$ are $L^2$
eigenvalues.

\subsection{Congruence: two cusps}

For $\p$ a prime ideal with $N(\p)=q$ and $\mathcal{O}/\p$ a field, the
quotient $\Gamma_0(\p)\backslash\HH$ has exactly two cusps $\infty$ and
$0$, with coset multiplicities $(1,q)$ under the identification
$\Gamma_0(\p)\backslash\Gamma\cong\mathbb{P}^1(\mathcal{O}/\p)$.

\begin{proposition}[Cusp cross-section areas]\label{prop:T0}
For $\p$ prime with $N(\p)=q$ and $\mathcal{O}/\p$ a field,
\[
  |T_\infty|=\frac{\sqrt{3}}{6},\qquad
  |T_0|=q\cdot|T_\infty|=\frac{q\sqrt{3}}{6}.
\]
In particular, for $\p=(1-\omega)$ one has $q=3$ and $|T_0|=\sqrt{3}/2$.
\end{proposition}

\begin{proof}[Sketch]
Cusp $\infty$: $\Gamma_\infty\cap\Gamma_0(\p)=\Gamma_\infty$, so the
cross-section equals the level-one section $|T|=\sqrt{3}/6$.  Cusp $0$:
the conjugated lattice is $\p$ (as an ideal), of covolume
$q\cdot\mathrm{covol}(\mathcal{O})$; after the order-$3$ unit fold the
orbifold area is $q\cdot|T|$.  Full writeup: companion note on $T_0$ areas.
\end{proof}

The multi-cusp zero-mode weights are therefore
\begin{equation}\label{eq:beta2}
  \beta_\infty(s)=\frac{1-s}{|T_\infty|Y^2},\qquad
  \beta_0(s)=\frac{1-s}{|T_0|Y^2}
  =\frac{1-s}{N(\p)|T_\infty|Y^2},
\end{equation}
with the same $\kappa_c(s)=1/((1+s)Y^2)$ for both cusps (equal truncation
heights).  The multi-cusp criterion is the natural analogue of Paper~II:
positivity of
$Q-\lambda M-\beta_\infty t_\infty^2-\beta_0 t_0^2$ on
$\{\mathcal{L}_s=0\}$ for all $s\in(0,1)$.

% ---------------------------------------------------------------------
\section{Guaranteed FEM lower bound}\label{sec:G1}

\subsection{Level one: Theorem G1}

The computational mesh is a tetrahedral mesh of $K_Y$ over the planar
domain $P_3$ (\texttt{build\_P3\_cell}): prism extrusion of a triangulation
of $P_3$, with base vertices ordered by index for conformity of shared
faces.  Floor lift (Lemma~G of Paper~I) produces an inner polyhedral mesh
$K_h\subset K_Y$; the sliver is absorbed by Lemma~S.  Crouzeix--Raviart
elements, the self-contained interpolation constant
$\kappa_{\mathrm{sc}}=\sqrt{1/\pi^2+1/15}$ (Lemma~I1 of Paper~I; inputs
Payne--Weinberger / Bebendorf only), and the exact-main-term refinement
are as in Paper~I.

\begin{theorem}[G1, level 1]\label{thm:G1}
On each $s$-window $W$, the continuum pencil inequality follows from
three finite checks: $c_\Sigma\ge 0$, $c_e>d_e$, and
$N_h-D_h\succeq 0$ on the CR space.  If these hold for a family of
windows covering $(0,1)$, then Theorem~\ref{thm:A} follows.
\end{theorem}

Full proofs of the coefficient formulas are in Paper~I (master theorem)
and the companion theory note; nothing in the chain is special to
$\ZZ[i]$ beyond the fundamental domain and $|T|$.

\subsection{Congruence: Theorem G1$\p$ and the reference-cell principle}

As in Paper~II, the congruence quotient is tiled by
$\mathrm{NC}=N(\p)+1$ isometric copies of the level-one reference cell,
glued by exact arithmetic over $\mathcal{O}/\p$.  Multi-copy rule:
\begin{quote}
copy $c$, $\delta$-source face
\;$\xrightarrow{\mathrm{glue}}$\;
copy $j=c\cdot\delta^{-1}$, $\delta$-target face,
\end{quote}
via the reference map $\varphi_\delta$ in pulled-back cell coordinates.
If $j=c$ the identification is internal and is relaxed (Neumann); if
$j\neq c$ the interface is interior and must be conforming.

\begin{lemma}[D0, tops free]\label{lem:D0}
When multi-copy top faces at $y=Y$ remain free (not in the pairing
dictionary), the CR interpolant reproduces both cusp trace functionals
exactly: $t_\infty(e)=t_0(e)=0$.  Consequently
\[
  \sigma_h=\sqrt{\mathrm{NC}}\,\alpha_h^{\mathrm{ref}}
\]
and the boundary blocks in $D_h$ are uninflated.
\end{lemma}

\begin{theorem}[G1$\p$]\label{thm:G1p}
With the multi-cusp criterion, the reference-cell assembly of
$\mathrm{NC}$ copies, Lemma~\ref{lem:D0}, and the coefficient list of
Paper~II (with $|T_\infty|=\sqrt{3}/6$ and $|T_0|=N(\p)|T_\infty|$ in the
$\beta$-weights), the finite checks $c_\Sigma\ge 0$, $c_e>d_e$, and
$N_h-D_h\succeq 0$ on a covering family of windows imply
Theorem~\ref{thm:B}.
\end{theorem}

\subsection{$N=3$ specialities}

For $\p=(1-\omega)$ one has $\mathcal{O}/\p\cong\mathbb{F}_3$ and
$\omega\equiv 1$ in $\mathbb{F}_3$.  Consequently $T_1\equiv T_\omega$ as
residue translations, and $U\equiv\mathrm{id}$ on $\mathbb{P}^1(\mathbb{F}_3)$.
Cross-copy glue uses $T_1/T_\omega$ and $S$; geometric $U$-maps remain
internal or residue-trivial.  Coset count: $|\mathbb{P}^1(\mathbb{F}_3)|=4$.
These simplifications are used by the production gluing code; they do not
affect the continuum criterion.

% ---------------------------------------------------------------------
\section{Verified certificate}\label{sec:cert}

\subsection{Arithmetic stack}

Both theorems use the same verification standard as Paper~I:
\begin{itemize}
\item Geometry and element matrices in Arb (python-flint, $128$-bit).
\item Scalar window inequalities by rigorous ball comparison.
\item Matrix positive-semidefiniteness by Rump \cite{Rump06}
  (Thm.~2.3, Cor.~2.4, Lemma~2.5), with power-of-two diagonal
  equilibration and per-row radii (Paper~I, Propositions on SAS and
  per-row absorption).
\item Element moments $Q$, $M$, $a$, $t$ by Taylor enclosures of order
  $p=5$ (weighted-moments routine shared with the Gaussian
  congruence certifier).
\end{itemize}

\subsection{Frozen parameters (Theorem~\ref{thm:A})}

\begin{table}[ht]
\centering
\begin{tabular}{@{}ll@{}}
\toprule
Parameter & Value \\
\midrule
Mesh & $P_3$, $N_{\mathrm{tri}}=6$, $N_3=3$
  $\to$ $648$ tets, $1440$ CR dofs \\
$Y$ & $1.25$ \\
$(\rho,\theta,\theta_2,\theta',\alpha)$ & $(55,\,0.7,\,0.5,\,0.5,\,0.5)$ \\
$\nu^*$ & $1.05$ \\
$s$-grid & $8$ uniform windows covering $[0,1]$ \\
Taylor $p$ / arb prec & $5$ / $128$ \\
$\kappa$ & $\kappa_{\mathrm{sc}}=\sqrt{1/\pi^2+1/15}$ \\
\bottomrule
\end{tabular}
\caption{Frozen certificate for Theorem~\ref{thm:A} (level 1).}
\label{tab:A}
\end{table}

Certified run (2026-07-11), abridged: Lemma~G admissible
($72$ floor faces); $1^TM_{\mathrm{mid}}1\approx 0.076661$
(matches $\vol(K_Y)$ up to the $O(h^2)$ sliver);
$t(\mathbf{1})=\sqrt{3}/6$; all eight windows: $c_e>d_e$ true, Rump PSD
true.  Conclusion: $P_3$/$\PSL(2,\ZZ[\omega])$ certified.

\subsection{Frozen parameters (Theorem~\ref{thm:B}, $N=3$)}

\begin{table}[ht]
\centering
\begin{tabular}{@{}ll@{}}
\toprule
Parameter & Value \\
\midrule
Reference mesh & same $P_3$ $6\times 3$ \\
Copies $\mathrm{NC}$ & $4$ \\
$Y$ & $1.25$ \\
$(\theta,\theta_2,\alpha,\theta_4,\tilde\rho)$
  & $(0.5,\,0.9,\,0.2,\,0.5,\,9)$ \\
$\nu^*$ & $1.05$ \\
$s$-grid & $8$ windows \\
$\sigma_h$ & $\sqrt{4}\,\alpha_h^{\mathrm{ref}}$ (Lemma~D0) \\
Global dofs & $n\approx 5404$ \\
Taylor $p$ / arb prec & $5$ / $128$ \\
\bottomrule
\end{tabular}
\caption{Frozen certificate for Theorem~\ref{thm:B}
($\Gamma_0(1-\omega)$).}
\label{tab:B}
\end{table}

Certified run (2026-07-12), abridged: Lemma~G arb OK;
$t_\infty(\mathbf{1})=|T_\infty|$, $t_0(\mathbf{1})=|T_0|$;
$\min c_e/d_e\gtrsim 11$; Rump PSD $8/8$.  Conclusion:
$\Gamma_0(1-\omega)$ certified.  Float precursor: min constrained
Rayleigh quotient $\mu\approx 1.12>0$.

\subsection{Frozen parameters (Theorem~\ref{thm:Btwo}, $N=7$)}

\begin{table}[ht]
\centering
\begin{tabular}{@{}ll@{}}
\toprule
Parameter & Value \\
\midrule
Reference mesh & same $P_3$ $6\times 3$ \\
Copies $\mathrm{NC}$ & $8$ \\
$Y$ & $1.25$ \\
$(\theta,\theta_2,\alpha,\theta_4,\tilde\rho)$
  & $(0.3,\,0.85,\,0.1,\,0.3,\,1)$ \\
$\nu^*$ & $1.001$ \\
$s$-grid & $8$ windows \\
$\sigma_h$ & $\sqrt{8}\,\alpha_h^{\mathrm{ref}}$ (Lemma~D0) \\
Global dofs & $n\approx 10533$ ($\sim 888$\,MB dense $A$) \\
Taylor $p$ / arb prec & $5$ / $128$ \\
\bottomrule
\end{tabular}
\caption{Frozen certificate for Theorem~\ref{thm:Btwo}
($\Gamma_0(\pi\mid 7)$).}
\label{tab:Btwo}
\end{table}

Certified run (2026-07-12): $t_\infty(\mathbf{1})=|T_\infty|$,
$t_0(\mathbf{1})=7\cdot|T_\infty|$; scalar $c_e>d_e$ $8/8$; Rump PSD
$8/8$.  Conclusion: $\Gamma_0(N=7)$ certified.  Float precursor on a
coarser mesh ($4\times 2$): $\min\mu\approx 1.12>0$.  Parameter search
used a float pencil LOBPCG screen (generalized $\min\mathrm{eig}(N_h,D_h)>1$)
before Rump, in the spirit of Paper~II.

\subsection{Congruence ingredients (checklist)}

\begin{center}
\begin{tabular}{@{}lll@{}}
\toprule
Ingredient & Artifact & Status \\
\midrule
Cosets $\cong\mathbb{P}^1(\mathcal{O}/\p)$
  & residue module & exact; $N=3$ has $4$ points \\
Gluing permutations (right action by $\delta^{-1}$)
  & same & $S^2=\mathrm{id}$, $U^3=\mathrm{id}$, round-trips \\
Cusp areas $|T_\infty|$, $|T_0|$
  & Prop.~\ref{prop:T0} & proved \\
Face dictionary on $P_3$
  & face-pairing module & RIGHT$\leftrightarrow$LEFT, etc. \\
Pairing matrices $g\in\Gamma$
  & Lemma~\ref{lem:pair} & frozen SL$(2,\mathcal{O})$ + checks \\
Lemma~D0 (tops free)
  & face dict excludes TOP & verified in cert \\
\bottomrule
\end{tabular}
\end{center}

% ---------------------------------------------------------------------
\section{Why each check suffices}

An eigenfunction $u$ with eigenvalue $\lambda=1-s^2\in(0,1)$ is $L^2$ and
real-analytic.  Cusp zero modes are exactly of the form $c(y/Y)^{1-s}$
(per cusp).  The Lax--Phillips reduction turns $Q(u)-\lambda M(u)=0$ into
$\mathcal{A}_s(u|_K)\le 0$ with $\mathcal{L}_s(u|_K)=0$, while
$u|_K\neq 0$ by unique continuation.  The pencil inequality, verified for
all $s$ through the finite window checks of Theorem~G1 / G1$\p$, gives
$\mathcal{A}_s>0$ on $\{\mathcal{L}_s=0\}$---contradiction.  The spectral
decomposition upgrades ``no eigenvalues in $(0,1)$'' to ``no spectrum in
$(0,1)$''.

% ---------------------------------------------------------------------
\section{Relation to other results in this series}

\begin{center}
\begin{tabular}{@{}lll@{}}
\toprule
Result & Group & Status \\
\midrule
Paper~I, main thm
  & $\PSL(2,\ZZ[i])$ level 1 & certified \\
Paper~II, Thms $N\p=5,9,13$
  & $\Gamma_0(\p)\le\PSL(2,\ZZ[i])$ & certified \\
\textbf{Theorem~\ref{thm:A}}
  & $\PSL(2,\ZZ[\omega])$ level 1 & \textbf{this paper} \\
\textbf{Theorem~\ref{thm:B}}
  & $\Gamma_0(1-\omega)$, $N=3$ & \textbf{this paper} \\
\textbf{Theorem~\ref{thm:Btwo}}
  & $\Gamma_0(\pi\mid 7)$, $N=7$ & \textbf{this paper} \\
STF / CE gate $B<1$ for $\omega$
  & same group, TF path & companion repo \\
Ladder rung $N=13$ over $\omega$
  & $\Gamma_0(\pi)$ & combinatorics only \\
\bottomrule
\end{tabular}
\end{center}

The STF/CE path and the FEM path are independent checks of the same
level-one spectral bound for $\PSL(2,\ZZ[\omega])$.
Theorems~\ref{thm:B}--\ref{thm:Btwo} have no STF counterpart in the
repository.

% ---------------------------------------------------------------------
\section{Dependencies and reproducibility}\label{sec:deps}

\subsection{External mathematical inputs}

\begin{enumerate}
\item \textbf{Spectral decomposition / continuous spectrum $\ge 1$}---
  Friedman \cite[Thm.~3.8.1]{Friedman} (cf.\ \cite[§6.2]{EGM98}).  Used
  for the $\lambda_1\ge 1$ phrasing; ``no $L^2$ eigenvalues in $(0,1)$''
  is the machine-certified core.
\item \textbf{EGM fundamental domain $F_d$}---
  Elstrodt--Grunewald--Mennicke \cite[§7.2--7.3]{EGM98}; planar domain
  $P_3$ as in \cite[§2.3]{DP20} (or equivalent Swan/EGM presentation).
\item \textbf{Shimizu horoball / precisely invariant cusp}---self-contained
  as in Paper~I (Shimizu \cite{Shimizu63}; Leutbecher \cite{Leutbecher67});
  same argument with $N(c)\ge 1$ for $\mathcal{O}=\ZZ[\omega]$.
\item \textbf{Real-analyticity of eigenfunctions / unique continuation}
  (Morrey--Nirenberg; elliptic regularity).
\item \textbf{Rump \cite{Rump06}}---Thm.~2.3 / Cor.~2.4 / Lemma~2.5.
\item \textbf{CR interpolation constant}---Lemma~I1 of Paper~I
  (Payne--Weinberger \cite{PW60} / Bebendorf \cite{Bebendorf03}); optional
  CZZ/CP sharpening not load-bearing.
\item \textbf{Side-pairing generators of $\Gamma$}---Swan; EGM Ch.~7;
  matrices and mesh dictionary frozen and machine-checked
  (Lemma~\ref{lem:pair}).
\end{enumerate}

Software correctness-critical stack: python-flint (Arb); IEEE-754 double
for Rump's floating Cholesky (as permitted by Rump: any LAPACK
\texttt{dpotrf}).

\subsection{Explicitly not used}

The Selberg trace formula; any test function; the scattering matrix or
Dedekind zeta of $K$; the torsion or conjugacy-class inventories; Then's
spectrum; the Gaussian ladder certificates of Paper~II (cited for method
only); and every result file of the companion trace-formula engine.

\subsection{Reproduce}

From \texttt{eisenstein\_numbers/}:
\begin{verbatim}
# Geometry freeze
python -u geometry_fund.py
python -u check_generators.py
python -u pairing_matrices.py
python -u non_neumann_omega.py 6 3   # optional paired FE

# Level 1 (Theorem A)
python -u cr_omega.py
python -u cert_omega.py 6 3

# Congruence rungs (Theorems B, B2)
python -u residue_omega.py
python -u cusps_omega.py
python -u face_pairings_p3.py
python -u congruence_omega_proto.py 3 6 3
python -u congruence_omega_proto.py 7 4 2
python -u cert_omega_p.py 3 6 3
python -u cert_omega_p.py 7 6 3
\end{verbatim}

% ---------------------------------------------------------------------
\section{What is not claimed}

\begin{itemize}
\item No claim for $\Gamma_0(\p)$ at $N(\p)=13$ (combinatorics
  only so far: residue rings, gluing permutations, and cusp-class
  partitions exist; interval certificates do not).
\item No re-proof of Theorems for $N\p=5,9,13$ on $\ZZ[i]$ (Paper~II).
\item Congruence multi-copy glues cross-copy faces via the frozen
  dictionary plus residue action; self-identifications remain Neumann.
\item The STF constant $B(\mathbb{Q}(\omega))$ is a separate software
  gate; the certificates of
  Theorems~\ref{thm:A}--\ref{thm:Btwo} do not use
  the trace formula or conjugacy-class inventory.
\end{itemize}

% ---------------------------------------------------------------------
\section{Outlook}

The natural next rung over $\ZZ[\omega]$ is the prime of norm $13$
(field residue $\mathbb{F}_{13}$), where the combinatorial infrastructure
already exists.  Certificate cost at $\mathrm{NC}=N+1$ copies grows as
$n\sim 10^4$ already at $N=7$; the Gaussian $N\p=13$ run of Paper~II
shows the same stack remains viable at $n\sim 3\cdot 10^4$ with tighter
$\nu^*$ and more $s$-windows.  The mode-bound ceiling and the
reference-cell principle are field-independent once $|T|$ and the
pairing dictionary are installed.

Together with Papers~I and~II, the series establishes a uniform
architecture---reduction, guaranteed FEM lower bound, verified
certificate---for excluding exceptional eigenvalues on Bianchi orbifolds
and their first congruence covers, for both class-number-one fields of
discriminants $-4$ and $-3$, by a method that never touches the Selberg
trace formula.

\begin{thebibliography}{99}

\bibitem{Bebendorf03}
M.~Bebendorf,
\emph{A note on the Poincar\'e inequality for convex domains},
Z.\ Anal.\ Anwend.\ \textbf{22} (2003), 751--756.

\bibitem{DP20}
C.~D\'oria and M.~Paula,
\emph{Height estimates for Bianchi groups},
arXiv:1910.03148.

\bibitem{EGM98}
J.~Elstrodt, F.~Grunewald, and J.~Mennicke,
\emph{Groups Acting on Hyperbolic Space},
Springer, 1998.

\bibitem{Friedman}
J.~Friedman,
\emph{The Selberg trace formula for $\mathrm{PSL}(2,O_K)$},
arXiv:math/0612807.

\bibitem{Leutbecher67}
A.~Leutbecher,
\emph{\"Uber die Heckeschen Gruppen $G(\lambda)$},
Math.\ Z.\ \textbf{100} (1967), 183--200.

\bibitem{PaperI}
\emph{Exceptional eigenvalues of Bianchi orbifolds by verified finite
elements.  I: The method, and the spectral gap of the Picard orbifold},
companion paper (this series).

\bibitem{PaperII}
\emph{Exceptional eigenvalues of Bianchi orbifolds by verified finite
elements.  II: Congruence levels of the Picard group},
companion paper (this series).

\bibitem{PW60}
L.~E.~Payne and H.~F.~Weinberger,
\emph{An optimal Poincar\'e inequality for convex domains},
Arch.\ Rational Mech.\ Anal.\ \textbf{5} (1960), 286--292.

\bibitem{Rump06}
S.~M.~Rump,
\emph{Verification of positive definiteness},
BIT Numer.\ Math.\ \textbf{46} (2006), 433--452.

\bibitem{Shimizu63}
H.~Shimizu,
\emph{On discontinuous groups operating on the product of the upper
half planes},
Ann.\ of Math.\ (2) \textbf{77} (1963), 33--71.

\end{thebibliography}

\end{document}
