% =====================================================================
% PAPER II ("The Strike")
% Exceptional eigenvalues of Bianchi orbifolds by verified finite
% elements. II: Congruence levels of the Picard group
% ---------------------------------------------------------------------
% Sources: independent_exclusion/{PROOF,CONGRUENCE,PROGRESS_LADDER,
%          HANDOFF_LADDER,0cuspchecks,theorem4postmortem}.md
%          m3p_certify.py, congruence_prototype.py
% =====================================================================
\documentclass[11pt]{amsart}
\usepackage{amsmath,amssymb,amsthm,booktabs}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{corollary}[theorem]{Corollary}
\newcommand{\HH}{\mathbb{H}^3}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\PSL}{\mathrm{PSL}}
\newcommand{\p}{\mathfrak{p}}
\newcommand{\vol}{\mathrm{vol}}

\title[Verified exclusion of exceptional eigenvalues, II]{Exceptional
eigenvalues of Bianchi orbifolds by verified finite elements.\\
II: Congruence levels of the Picard group}
\author{}
\date{Draft, 2026-07-11}

\begin{document}
\begin{abstract}
Using the verified finite-element exclusion machinery of Paper~I, we
prove that the Hecke congruence quotients
$\Gamma_0(\p)\backslash\HH$ of the Picard group, for the Gaussian
primes $\p$ of norm $5$, $9$, and $13$, have no Laplace eigenvalue in
$(0,1)$.  These quotients lie beyond the reach of trace-formula
positivity, whose margin degrades linearly with volume.  The key
structural input is a reference-cell principle: the congruence quotient
is tiled by exactly $N\p+1$ isometric copies of the level-one cell,
glued by exact arithmetic over $\ZZ[i]/\p$, with both cusp collars
assembling as products at the same truncation height---so every
interval-arithmetic error bound is inherited from level one unchanged,
and only the size of the final verified linear algebra grows with the
index.  All three certificates are deterministic and reproducible.
\end{abstract}
\maketitle

% ---------------------------------------------------------------------
\section{Introduction}

Paper~I \cite{PaperI} developed a machine-verified method for excluding
Laplace eigenvalues in $(0,1)$ on cusped hyperbolic $3$-orbifolds:
Lax--Phillips cusp reduction, guaranteed Crouzeix--Raviart lower bounds,
and a Rump positive-definiteness certificate, with no use of the Selberg
trace formula.  As validation it proved $\lambda_1\ge 1$ for the Picard
orbifold $\PSL(2,\ZZ[i])\backslash\HH$---a statement believed classically
known, and independently certified by a trace-formula positivity argument
in the companion repository ($B(\ZZ[i])\le 0.3199<1$).

The present paper takes the method into open territory.  Let
$\Gamma=\PSL(2,\ZZ[i])$ and let $\p\subset\ZZ[i]$ be a Gaussian prime
ideal.  The Hecke congruence subgroup
\[
  \Gamma_0(\p)
  =\Bigl\{\begin{pmatrix}a&b\\c&d\end{pmatrix}\in\Gamma:\;c\in\p\Bigr\}
\]
has index $N\p+1$ in $\Gamma$ (for $\p$ prime), two cusps when $\p$ is
prime, and covolume $(N\p+1)\cdot\vol(\Gamma\backslash\HH)$.  Trace-formula
positivity degrades linearly with volume: already at $N\p=5$ one has
$B\approx 1.9>1$, so the $B<1$ gate fails outright.  The finite-element
margin, by contrast, tracks the actual spectral gap and degrades only
sublinearly.  The measured ladder of float margins is
\begin{equation}\label{eq:ladder}
  \mu\approx 7.3,\; 4.4,\; 2.4,\; 1.45
  \qquad\text{for }N\p=1,\,5,\,9,\,13
\end{equation}
(the level-one value from Paper~I; the rest measured on the glued CR
space as constrained Rayleigh quotients).  All four lie comfortably above
the certification floor near $1$.

\begin{theorem}[$N\p=5$]\label{thm:5}
Let $\p=(2+i)$.  The Laplacian on $L^2(\Gamma_0(\p)\backslash\HH)$ has
no eigenvalue in $(0,1)$.
\end{theorem}

\begin{theorem}[$N\p=9$]\label{thm:9}
Let $\p=(3)$ (inert).  The Laplacian on $L^2(\Gamma_0(\p)\backslash\HH)$
has no eigenvalue in $(0,1)$.
\end{theorem}

\begin{theorem}[$N\p=13$]\label{thm:13}
Let $\p=(3+2i)$ (split).  The Laplacian on
$L^2(\Gamma_0(\p)\backslash\HH)$ has no eigenvalue in $(0,1)$.
\end{theorem}

The computational engine---Theorem~G1, Lemma~I1, Arb enclosures, Rump's
certificate, including the two refinements of Paper~I (error-free
diagonal congruence and per-row radius absorption)---is cited from
Paper~I and not re-derived.  What is genuinely new here is:
\begin{enumerate}
\item Facts~A--B (exact tiling by isometric copies; collar assembly at
  equal height);
\item the conforming/relaxed gluing dichotomy;
\item the multi-cusp criterion and Theorem~G1$\p$ (the two-cusp
  bookkeeping delta);
\item the residue-arithmetic gluing for $\mathbb{F}_5$, $\mathbb{F}_9$,
  and $\mathbb{F}_{13}$;
\item the frozen certificates and the $N\p=13$ stress test that forced
  the Rump refinements into load-bearing status.
\end{enumerate}
Reviewer burden is confined to (i)~the reference-cell principle,
(ii)~the exact coset gluing, and (iii)~the multi-cusp geometry.  Nothing
numerical-analytic is new except the per-cusp bookkeeping of
Theorem~G1$\p$.

Table~\ref{tab:ladder} summarizes the three rungs against level one.

\begin{table}[ht]
\centering
\begin{tabular}{@{}lcccc@{}}
\toprule
$N\p$ & Index & $\vol$ (approx.) & Float $\mu$ & TF $B$ (approx.) \\
\midrule
$1$ (level 1) & $1$ & $0.305$ & $7.3$ & $0.32$ \\
$5$ & $6$ & $1.83$ & $4.4$ & $1.9$ \\
$9$ & $10$ & $3.05$ & $2.4$ & $\gg 1$ \\
$13$ & $14$ & $4.27$ & $1.45$ & $\gg 1$ \\
\bottomrule
\end{tabular}
\caption{Margin ladder.  Trace-formula $B$ scales roughly as volume;
the FE margin $\mu$ degrades sublinearly.  Rungs $N\p\ge 5$ are
unreachable by $B<1$ positivity.}
\label{tab:ladder}
\end{table}

% ---------------------------------------------------------------------
\section{The multi-cusp criterion}\label{sec:crit}

\subsection{Cusps of $\Gamma_0(\p)$}

For $\p$ prime the quotient $\Gamma_0(\p)\backslash\HH$ has exactly two
cusps, represented by $\infty$ and $0$.  Scaling matrices
$\sigma_\alpha\in\Gamma$ conjugate the stabilizers; for cusp $0$ one may
take $\sigma_0=\bigl(\begin{smallmatrix}0&-1\\1&0\end{smallmatrix}\bigr)$.
The conjugated stabilizer is
\[
  \sigma_0^{-1}
  \begin{pmatrix}a&b\\c&d\end{pmatrix}
  \sigma_0
  =\begin{pmatrix}d&-c\\0&a\end{pmatrix}
  \qquad\bigl(c\in\p\bigr),
\]
so the cusp-$0$ lattice is $\Lambda_0=\p$ (as an ideal in $\ZZ[i]$), and
the order-$2$ fold $z\mapsto -z$ is present at both cusps, exactly as at
level one.

\begin{table}[ht]
\centering
\begin{tabular}{@{}lllll@{}}
\toprule
Cusp $\alpha$ & Lattice $\Lambda_\alpha$ & $|T_\alpha|$
  & Shortest dual $|\mu|$ & Mode bound \\
\midrule
$\infty$ & $\ZZ[i]$ & $1/2$ & $1$
  & $4\pi^2 Y^2\approx 61.7$ \\
$0$ & $\p$ & $N\p/2$ & $1/\sqrt{N\p}$
  & $4\pi^2 Y^2/N\p$ \\
\bottomrule
\end{tabular}
\caption{Per-cusp data at truncation height $Y=1.25$.  Mode bounds must
exceed $\lambda<1$; the inequality $N\p<4\pi^2 Y^2\approx 61.7$ holds
through the entire near-term ladder.}
\label{tab:cuspdata}
\end{table}

\subsection{Zero-mode defects and the constraint}

Write $t_\alpha(v)=\int_{T_\alpha}v(\cdot,Y)\,dx$ for the Euclidean trace
integral on the cusp-$\alpha$ cross-section (so
$c_\alpha=t_\alpha/|T_\alpha|$ is the zero-mode coefficient).  The
level-one Lemmas~1--3 of Paper~I apply verbatim per cusp:

\begin{itemize}
\item \textbf{Zero-mode defect.}  An eigenfunction with
  $\lambda=1-s^2$ contributes
  \[
    -\sum_\alpha\frac{|T_\alpha|(1-s)c_\alpha^2}{Y^2}
    =-\frac{1-s}{Y^2}\Bigl(2\,t_\infty^2+\frac{2}{N\p}\,t_0^2\Bigr)
  \]
  to $Q-\lambda M$ from the two collars.  Hence
  \[
    \beta_\infty(s)=\frac{2(1-s)}{Y^2},\qquad
    \beta_0(s)=\frac{2(1-s)}{N\p\,Y^2}.
  \]
\item \textbf{Nonzero modes.}  Nonnegative by the mode bounds of
  Table~\ref{tab:cuspdata} (both $\gg 1$ for $N\p\in\{5,9,13\}$).
\item \textbf{Orthogonality to constants.}  With equal truncation heights
  $Y_\infty=Y_0=Y$,
  \[
    \mathcal{L}_s(v)=a(v)+\kappa_c(s)\bigl(t_\infty(v)+t_0(v)\bigr),\qquad
    \kappa_c(s)=\frac{1}{(1+s)Y^2}.
  \]
  The constraint stays rank-one in the combined vector $t_\infty+t_0$.
\end{itemize}

\begin{theorem}[Multi-cusp criterion]\label{thm:multicrit}
If for every $s\in(0,1)$ and every $0\neq v\in H^1(D_Y)$ with
$\mathcal{L}_s(v)=0$,
\[
  Q(v)-(1-s^2)M(v)-\beta_\infty(s)\,t_\infty(v)^2
  -\beta_0(s)\,t_0(v)^2>0,
\]
then $\Gamma_0(\p)\backslash\HH$ has no eigenvalue in $(0,1)$.
\end{theorem}

The pencil form is identical to Paper~I's Lemma~P, with the rank-one
boundary term replaced by a rank-two form
$\beta_\infty t_\infty^2+\beta_0 t_0^2$.  Precise invariance of both
collars is inherited from level one: $\{y>1\}$ is precisely invariant
under the full group $\Gamma$ (Paper~I, Shimizu lemma), both scaling
matrices $\sigma_\alpha$ lie in $\Gamma$, and $\Gamma_0(\p)\subset\Gamma$
needs nothing new.

% ---------------------------------------------------------------------
\section{The reference-cell principle}\label{sec:refcell}

This section is the conceptual heart of the paper.  It explains why
interval-arithmetic cost does not grow with the index.

\begin{proposition}[Fact A: exact tiling]\label{prop:A}
Right cosets of $\Gamma_0(\p)$ in $\Gamma$ are in bijection with
$\mathbb{P}^1(\ZZ[i]/\p)$.  If $\{\gamma_i\}$ is a set of representatives and
$F$ is the Humbert fundamental domain, then
$D=\bigcup_i\gamma_i F$ is a fundamental domain for $\Gamma_0(\p)$, and
each copy $\gamma_i F$ is isometric to $F$.  Assembling each copy in its
own pulled-back reference coordinates makes all per-copy element data
equal to the reference data; only the boundary identifications know the
embedding.
\end{proposition}

\begin{proposition}[Fact B: collar assembly]\label{prop:B}
Truncate every copy at its reference height $Y$.  The cusp-$\infty$
collar is $T\times(Y,\infty)$ exactly as at level one.  For cusp $0$,
take $\sigma_0=\bigl(\begin{smallmatrix}0&-1\\1&0\end{smallmatrix}\bigr)$
and representatives $\gamma_k=\bigl(\begin{smallmatrix}0&-1\\1&k
\end{smallmatrix}\bigr)$ for $k$ running over lifts of $\ZZ[i]/\p$.  Then
\[
  \sigma_0^{-1}\gamma_k=\begin{pmatrix}1&k\\0&1\end{pmatrix},
\]
a \emph{height-preserving} translation.  In cusp-$0$ normalized
coordinates the chimneys $\gamma_k(\{y>Y\}\times T)$ are
$(T+k)\times(Y,\infty)$, and the translates $T+k$ tile the cusp-$0$
cross-section $T_0$ (fundamental domain of
$\p\rtimes\langle -1\rangle$, area $N\p/2$).  Fold-compatibility of the
tiling $T_0=\bigsqcup_k(T+k)$ uses the coset identity
$k\equiv\varepsilon\lambda\pmod\p$ under the order-$2$ action---the one
step that is not automatic in the semidirect product (audited as item~C2
in the companion cusp checklist).
\end{proposition}

Consequently the assembled truncated domain is exactly
\[
  D_Y
  =\bigl[(N\p+1)\text{ isometric copies of }K\bigr]
  \;\cup\;
  T\times(Y,\infty)
  \;\cup\;
  T_0\times(Y,\infty).
\]
Volume sanity check:
\[
  (N\p+1)\Bigl(\vol(F)-\frac{1}{4Y^2}\Bigr)
  +\frac{\tfrac12+N\p/2}{2Y^2}
  =(N\p+1)\,\vol(F).
\]

\begin{proposition}[Reference-cell principle]\label{prop:RCP}
In the level-$\p$ certificate, every interval-enclosed quantity is either
\begin{enumerate}
\item[(a)] a per-cell quantity of the level-one reference cell $K$
  (element matrices, $\ell_0$ Taylor enclosures, Lemmas~E/G/S constants---
  computed once, radii independent of the index), or
\item[(b)] exact integer combinatorics (coset action, gluing table, cusp
  classes).
\end{enumerate}
The only computation that grows with the index is the verified
positive-definiteness check (Rump), whose cost is linear algebra, not
error.
\end{proposition}

This is the scale protection that makes the ladder feasible: error
radii stay at level-one size; only $n\sim(N\p+1)\cdot n_{\mathrm{ref}}$
grows.

% ---------------------------------------------------------------------
\section{Exact gluing and the symmetric mesh}\label{sec:glue}

\subsection{Side pairings and the gluing rule}

Reference-cell side pairings of the Humbert domain $F$ and their face
maps:

\begin{center}
\begin{tabular}{@{}llll@{}}
\toprule
$\delta$ & Matrix & Faces & Reference face map \\
\midrule
$T_1$ & $(1,1;0,1)$ & $x_1=-\tfrac12\to x_1=+\tfrac12$
  & identity in $(x_2,y)$ \\
$R$ & $(i,0;0,-i)$ & $x_2=0\to$ itself & $x_1\mapsto -x_1$ \\
$T_i R$ & $(1,i;0,1)\cdot R$ & $x_2=\tfrac12\to$ itself
  & $x_1\mapsto -x_1$ \\
$S$ & $(0,-1;1,0)$ & floor $\to$ itself & $(x_1,x_2)\mapsto(-x_1,x_2)$ \\
\bottomrule
\end{tabular}
\end{center}

Cosets are labelled by bottom rows $(c:d)\in\mathbb{P}^1(\ZZ[i]/\p)$; the
right action is $(c:d)\cdot M$.  Copy $c$'s $\delta$-source face glues to
copy $j:=(c:d)\cdot\delta^{-1}$'s $\delta$-target face via the reference
face map.  Derivation: $\gamma_c x\sim\gamma_j(\delta x)$ if and only if
$\gamma_c\delta^{-1}\gamma_j^{-1}\in\Gamma_0(\p)$.

\subsection{The conforming/relaxed dichotomy}

\begin{itemize}
\item If $j=c$, the identification is internal to one copy and is
  \textbf{relaxed} (Neumann), exactly as at level one---valid, since it
  enlarges the test space.
\item If $j\neq c$, the interface is \emph{interior to the domain} and
  must be \textbf{conforming}: relaxing it would hand each copy its own
  constant function, and the criterion would (correctly) fail.
\end{itemize}
Cusp class of coset $(c:d)$: $\infty$ if and only if $c\equiv 0$; else
$0$.  Class sizes are $\{1,N\p\}$.

\subsection{The $24$-split mesh}

Conformity across interfaces requires the reference boundary
triangulation to be compatible with the face maps: identity/translation
across the $x_1$-faces \emph{and} the $x_1$-mirror on $x_2=0$,
$x_2=\tfrac12$, and the floor.  Kuhn splits (uniform diagonals) violate
the mirror maps.  Resolution: the \textbf{$24$-split}---each mapped hex
gets a body-centre node and six face-centre nodes; each quad face splits
into four triangles; each triangle plus body centre yields a tet.  This
split is invariant under all face maps in the table (the four-way face
split has no diagonal convention) and is automatically face-to-face
across translated neighbors.  Cost: four times the tets of Kuhn,
compensated by a coarser base grid.  Lemma~G (floor lift) reproves
verbatim on the quarter-triangles; Lemmas~E/S/I1/T and Theorem~G1 of
Paper~I are mesh-split-agnostic.  The base grid needs $N_1$ even
(mirror-symmetric grid lines).

\subsection{Residue arithmetic}

Three residue rings, one per theorem:

\begin{center}
\begin{tabular}{@{}llll@{}}
\toprule
Level & Ring & $N\p$ & $i$-image \\
\midrule
$(2+i)$ & $\mathbb{F}_5$ (prime field) & $5$ & $3$ \\
$(3)$ & $\mathbb{F}_9$ as pairs $(a,b)\bmod 3$, $i^2=-1$
  & $9$ & $(0,1)$ \\
$(3+2i)$ & $\mathbb{F}_{13}$ (prime field) & $13$ & $8$ \\
\bottomrule
\end{tabular}
\end{center}

The inert case $\ZZ[i]/(3)\cong\mathbb{F}_9$ is \emph{not} a prime field;
generic ring operations on pairs replace mod-$5$ integer tricks.  The
assert suite---permutations, round-trip action, involutions
$M^2\equiv\pm I$ in $\PSL$ for $R$, $T_i R$, $S$, class sizes
$\{1,N\p\}$, connectivity of the assembled dof graph,
$t_\infty(\mathbf{1})=\tfrac12$, $t_0(\mathbf{1})=N\p/2$, volume
identity---is the combinatorial error net.  The generalized $\mathbb{F}_5$
path is bit-identical to the original hardcode (regression check).

% ---------------------------------------------------------------------
\section{The guaranteed lower bound at level $\p$}\label{sec:G1p}

Everything is as in Paper~I's Theorem~G1 (whose lemmas apply per
reference cell verbatim) with the following changes.  We state only the
deltas.

\subsection{Lemma D0 per cusp}

The cusp cross-sections at height $Y$ are exactly tiled by boundary faces
of the mesh (top faces of the top-layer tets, per copy), and the CR
interpolant matches the mean of $v$ on every mesh face.  Hence, with
$e=v-I_{\mathrm{CR}}v$,
\[
  t_\infty(e)=t_0(e)=0:
\]
both trace functionals are exactly reproduced.  Consequently there are
no slab constants, no Young split of $t_\alpha(v)^2$ (boundary terms enter
$D_h$ with coefficients $\tilde\beta_\alpha$, uninflated), and
\[
  |a(e)|\le\sqrt{N\p+1}\,\alpha_h^{\mathrm{ref}}\sqrt{Q_{\mathrm{pc}}(e)}
\]
(disjoint copies), so
\[
  \sigma_h=\sqrt{N\p+1}\,\alpha_h^{\mathrm{ref}}
\]
and $d_e$ has no trace part.

\subsection{Multi-sliver absorption and window data}

Reference constants $\gamma$, $\alpha_h^{\mathrm{ref}}$, $S_M$, $S_Q$,
$S_\Sigma$, $V_\Sigma$ are computed once on the $24$-split reference mesh
(in arb).  The $N\p+1$ per-copy slivers are absorbed by the reference
Lemma~S constants (per-copy maxima; disjoint columns).  For a window
$W=[s^-,s^+]\ni s_0$ with $\lambda^+=1-(s^-)^2$,
$\beta_\infty^+=2(1-s^-)/Y^2$, $\beta_0^+=\beta_\infty^+/N\p$, and
$\Delta\kappa=\max_W|\kappa_c(s)-\kappa_c(s_0)|$, define
\begin{align*}
  d_2&:=\rho(1/\theta_2-1)=\tilde\rho/\theta_2,\qquad
  \tilde\rho:=\rho(1-\theta_2),\\
  \omega&:=2(N\p+1)\,d_2 V_\Sigma,\\
  c_Q&:=1-(\omega+\nu^*\lambda^+)S_Q,\qquad
  c_\Sigma:=1-(\omega+\nu^*\lambda^+)S_\Sigma,\\
  \tilde\lambda&:=\nu^*\lambda^+(1+S_M)+\omega S_M,\\
  \tilde\beta_\alpha&:=\nu^*\beta_\alpha^++4\,d_2\Delta\kappa^2,\\
  \sigma_h&:=\sqrt{N\p+1}\,\alpha_h^{\mathrm{ref}},\\
  c_e&:=c_Q\bigl(1-(1/\theta_4-1)\rho_w\bigr)
    -\tilde\rho(1/\theta-1)\sigma_h^2,\\
  d_e&:=\tilde\lambda(1+1/\alpha)\gamma^2.
\end{align*}
(The $\mathcal{L}$-shift error uses
$E^2\le 2(N\p+1)V_\Sigma M_\Sigma+4\Delta\kappa^2(t_\infty^2+t_0^2)$,
whence the factors in $\omega$ and $\tilde\beta$.)

\subsection{Exact main terms and discrete data}

As in Paper~I: exact-weight mass $M_{\mathrm{ex}}$, stiffness remainder
$Q_{\mathrm{rem}}\succeq 0$, and $\rho_w=\max_T(y_T^+/y_T^--1)$.  Discrete
data on the glued global CR space (exact residue scatter of the
reference element enclosures): $Q_{\mathrm{pc}}$, $Q_{\mathrm{rem}}$,
$M_{\mathrm{ex}}$, $a$, $t_\infty$, $t_0$,
$\ell_0=a+\kappa_c(s_0)(t_\infty+t_0)$;
\begin{align*}
  N_h&:=c_Q\bigl[Q_{\mathrm{pc}}+(1-\theta_4)Q_{\mathrm{rem}}\bigr]
    +\tilde\rho(1-\theta)\,\ell_0\ell_0^T,\\
  D_h&:=\tilde\lambda(1+\alpha)M_{\mathrm{ex}}
    +\tilde\beta_\infty\,t_\infty t_\infty^T
    +\tilde\beta_0\,t_0 t_0^T,
\end{align*}
with boundary terms uninflated by Lemma~D0.

\begin{theorem}[G1$\p$]\label{thm:G1p}
If $c_\Sigma\ge 0$, $c_e\ge d_e$, and $N_h-D_h\succeq 0$ on a family of
windows covering $(0,1)$, then $\Gamma_0(\p)\backslash\HH$ has no
eigenvalue in $(0,1)$.
\end{theorem}

\begin{proof}[Proof by delta]
Identical to Paper~I's Theorem~G1 with: the multi-cusp criterion of
Section~\ref{sec:crit} in place of the level-one criterion; the $N\p+1$
per-copy slivers absorbed by the reference Lemma~S constants; the window
shift of $\mathcal{L}$ carried out with the combined error $E$ above; the
CR projection step using Lemma~D0 for the per-cusp trace and volume
functionals; Lemma~R of Paper~I applied per vector (rank one positive for
$\ell_0$, two rank-one negatives for $t_\infty$, $t_0$) in the
interval-to-midpoint reduction; and the Rump certificate for the matrix
inequality.  Testing over all of $H^1$ of the glued core (conforming
across distinct-copy interfaces, free on self-identifications) is a
relaxation of the true form domain, as before.
\end{proof}

% ---------------------------------------------------------------------
\section{Certificates and results}\label{sec:results}

\subsection{Frozen certificates}

\begin{table}[ht]
\centering
\small
\begin{tabular}{@{}lccc@{}}
\toprule
 & Thm~\ref{thm:5} ($N=5$)
 & Thm~\ref{thm:9} ($N=9$)
 & Thm~\ref{thm:13} ($N=13$) \\
\midrule
Reference mesh & $8\times 4\times 3$
  & $6\times 3\times 3$ & $8\times 3\times 2$ \\
Tets/copy (24-split) & $2304$ & $1296$ & $1152$ \\
Global CR dofs & $28\,400$ & $26\,532$ & $33\,176$ \\
Copies $N\p+1$ & $6$ & $10$ & $14$ \\
$Y$ & $1.25$ & $1.25$ & $1.25$ \\
$\nu^*$ & $1.02$ & $1.02$ & $1.001$ \\
Windows & $8$ & $8$ & $16$ \\
$(\theta,\theta_2,\alpha,\theta_4,\tilde\rho)$
  & $(0.6,0.9,0.2,0.85,9.0)$
  & $(0.5,0.9,0.15,0.8,1.5)$
  & $(0.3,0.98,0.1,0.85,0.5)$ \\
$\kappa$ & $\kappa_{\mathrm{sc}}$ & $\kappa_{\mathrm{sc}}$
  & $\kappa_{\mathrm{sc}}$ \\
Taylor $p$ / arb prec & $5$ / $128$ & $5$ / $128$ & $5$ / $128$ \\
$c_e$ range & $\ge 0.847$
  & $[0.836,0.881]$ & $[0.870,0.907]$ \\
$d_e$ range & $\le 0.138$
  & $[0.079,0.222]$ & $[0.074,0.485]$ \\
Wall time (approx.) & --- & $38$ min & $93$ min \\
Certified & 2026-07-10 & 2026-07-11 & 2026-07-11 \\
\bottomrule
\end{tabular}
\caption{Frozen certificates for Theorems~\ref{thm:5}--\ref{thm:13}.
Every window: $c_\Sigma>0$, $c_e>d_e$, Rump PSD true.}
\label{tab:certs}
\end{table}

All three certificates use $\kappa=\kappa_{\mathrm{sc}}$ (Paper~I,
Lemma~I1), exact main terms, and Lemma~D0.  The parameter-search
methodology is: scalar conditions plus a constant-direction heuristic
plus a float pencil filter; floats choose, arb verifies.  A notable
structural feature is the scalar-slack / pencil anti-correlation through
$\tilde\rho$: high $\tilde\rho$ helps the constant direction but can
starve the pencil; the search balances both.

Observe that $\tilde\rho$ shrinks with $N\p$ ($9.0\to 1.5\to 0.5$): the
constant-direction scaling \emph{improves} with $N\p$, because $D(1)$'s
boundary part grows like $N\p/2$ while $\mathcal{L}(1)^2$ grows like
$N\p^2$, so the $\tilde\rho$ demand shrinks.

\subsection{The $N\p=13$ stress test}

At $N\p=13$ the float margin collapses to $\mu\approx 1.45$
(mesh-converged).  The continuum problem is still safely positive, but
the certificate becomes the binding constraint.  Specifically: the
midpoint matrix $A$ of $N_h-D_h$ was SPD with
$\lambda_{\min}(A)\approx 3\cdot 10^{-5}$, yet a uniform max-row-sum radius
shift of order $1.5\cdot 10^{-4}$ exceeded that margin.  The matrix was
SPD; the unscaled certificate failed.

Resolution by Paper~I's two refinements:
\begin{enumerate}
\item error-free power-of-two diagonal congruence
  $A\mapsto SAS$, $s_i=2^{\mathrm{round}(-\frac12\log_2 a_{ii})}$;
\item per-row radius absorption $r_i=\sum_j|S\,\Delta A\,S|_{ij}$
  subtracted from diagonal $i$.
\end{enumerate}
With these, all $16$ windows pass.  No change to Theorem~\ref{thm:G1p}.
This is the strongest evidence that the certificate---not the
spectrum---was the binding constraint at $N\p=13$, and that the machinery
scales: the same continuum theory, with only implementation hardening of
the linear-algebra step, reaches the third congruence rung.

\subsection{Margin ladder and scaling}

The ladder \eqref{eq:ladder} is sublinear in volume.  Trace-formula $B$
scales roughly as volume and is already $>1$ at the first congruence
rung.  Certificate cost is dominated by the dense Cholesky of an
$n\times n$ matrix with $n\sim(N\p+1)\,n_{\mathrm{ref}}$; at $N\p=13$ one
has $n\approx 3.3\cdot 10^4$ and wall time about $93$ minutes on a
workstation.  The mode-bound ceiling $N\p<4\pi^2 Y^2\approx 61$ leaves
room for several further prime levels before a taller $Y$ or more
windows become necessary.

% ---------------------------------------------------------------------
\section{Discussion and outlook}

\subsection{Where the ladder ends as-is}

Three constraints bound further progress without new ideas:
\begin{enumerate}
\item \textbf{Mode bound.}  Nonzero Fourier modes in the wider cusp
  remain nonnegative only while $N\p<4\pi^2 Y^2\approx 61$.  Past this,
  either raise $Y$ (enlarging the core and the mesh) or accept a more
  refined mode analysis.
\item \textbf{Certificate cost.}  Dense verified Cholesky is
  $O(n^3)$ with $n\sim(N\p+1)\,n_{\mathrm{ref}}$.  Out-of-core blocked
  Cholesky is admissible under Rump's theorem (``any library routine''),
  but is engineering rather than mathematics.
\item \textbf{Margin.}  Extrapolating the sublinear ladder, $\mu$ will
  eventually approach $1$; then $\nu^*\downarrow 1$ and more windows are
  needed, and eventually the continuum gap itself may sit below the
  discrete-check floor.
\end{enumerate}

\subsection{Composite levels and other fields}

Composite or split levels (e.g.\ $(5)=(2+i)(2-i)$) have more than two
cusps.  Facts~A--B extend with general cusp-width bookkeeping (still
finitely many height-preserving translation classes, but widths vary per
cusp).  Other Bianchi fields---notably $\ZZ[\omega]$, treated at level one
and at $\Gamma_0(1-\omega)$ in Paper~III---carry the same architecture
with a different fundamental domain and residue arithmetic.  Torsion-free
principal congruence covers $\Gamma(\mathfrak{n})$ eliminate the
orbifold structure entirely and are natural further targets.

\subsection{Relation to the classical program}

Certified gaps at small level do not prove the Selberg-type conjecture
over $\mathbb{Q}(i)$ in general.  They do establish the first few
congruence cases by a method whose margin tracks the gap rather than the
volume, and they supply independent validation of the level-one statement
already known by the trace formula.  The combination---exclusion on
$(0,1)$ by FEM, existence of the first Maass form by rigorous computation---
gives a rigorous picture of the bottom of the spectrum for every quotient
within reach of both methods.

% ---------------------------------------------------------------------
\section{Dependencies and reproducibility}\label{sec:deps}

\subsection{Dependency delta versus Paper~I: none}

The external-input list of Paper~I is unchanged.  The additional
ingredients---coset combinatorics, Facts~A--B, Lemma~D0 in multi-cusp
form, the cusp-$0$ audit (conjugated stabilizer, fold-compatible tiling,
height-preserving chimney maps, dual-lattice mode bound, inherited
precise invariance)---are all proved in-repo.  Friedman
\cite[Thm.~3.8.1]{Friedman} already covers cofinite Kleinian groups,
hence $\Gamma_0(\p)$ with trivial character verbatim.

Explicitly not used: the Selberg trace formula, test functions, scattering
data, torsion inventories, Then's spectrum, and every result file of the
companion trace-formula engine.

\subsection{Reproduce}

From \texttt{independent\_exclusion/}:
\begin{verbatim}
# Gluing asserts for all three levels
python congruence_prototype.py gluing

# Theorem 2 (N=5): default certify
python m3p_certify.py

# Theorem 3 (N=9)
python -u -c "from m3p_certify import certify; \
  certify(6,3,3,params=(0.5,0.9,0.15,0.8,1.5),level='(3)')"

# Theorem 4 (N=13)
python -u -c "import m3p_certify as m; m.NU_STAR=1.001; m.NWIN=16; \
  m.certify(8,3,2,params=(0.3,0.98,0.1,0.85,0.5),level='(3+2i)')"

# Float margins (diagnostic)
python congruence_prototype.py margin "(2+i)"
python congruence_prototype.py margin "(3)"
python congruence_prototype.py margin "(3+2i)"
\end{verbatim}

Software stack: python-flint (Arb), NumPy/SciPy/LAPACK \texttt{dpotrf},
IEEE-754 double.  Wall times: about $38$ minutes ($N\p=9$) and $93$
minutes ($N\p=13$) on a standard workstation for the certified runs of
2026-07-11.

% ---------------------------------------------------------------------
\appendix
\section{The cusp-$0$ audit}

Every cusp-$0$ claim used above admits a one-paragraph proof and a
numerical cross-check that would catch the corresponding error class.
We record the checklist (full writeups in the companion note
\texttt{0cuspchecks.md}).

\begin{enumerate}
\item[\textbf{C1.}] \emph{Conjugated stabilizer.}
  $\sigma_0^{-1}\bigl(\begin{smallmatrix}a&b\\c&d\end{smallmatrix}\bigr)
  \sigma_0=\bigl(\begin{smallmatrix}d&-c\\0&a\end{smallmatrix}\bigr)$
  for $c\in\p$; lattice $\Lambda_0=\p$; fold present.  Check:
  $t_0(\mathbf{1})=N\p/2$.
\item[\textbf{C2.}] \emph{Fold-compatible tiling.}
  The identity $k\equiv\varepsilon\lambda\pmod\p$ ensures
  $T_0=\bigsqcup_k(T+k)$ is a fundamental domain for
  $\p\rtimes\langle -1\rangle$.  Check: area sum $N\p\cdot|T|=N\p/2$.
\item[\textbf{C3.}] \emph{Height-preserving chimney maps.}
  $\sigma_0^{-1}\gamma_k=(1,k;0,1)$ preserves hyperbolic height.  Check:
  all class-$0$ copies contribute top faces at the same $y=Y$.
\item[\textbf{C4.}] \emph{Dual-lattice mode bound.}
  Shortest dual vector of $\Lambda_0$ has length $1/\sqrt{N\p}$; mode
  bound $4\pi^2 Y^2/N\p>1$ for $N\p<61.7$.  Check: tabulated for
  $N\p\in\{5,9,13\}$.
\item[\textbf{C5.}] \emph{Shared $\kappa_c$.}
  Equal truncation heights $\Rightarrow$
  $\kappa_c=1/((1+s)Y^2)$ for both cusps; constraint rank one in
  $t_\infty+t_0$.  Check: code builds a single $\ell_0$ vector.
\item[\textbf{C6.}] \emph{Inherited precise invariance.}
  Both horoballs are $\Gamma$-images of the level-one horoball;
  $\Gamma_0(\p)\subset\Gamma$.  Check: no $\Gamma_0$-specific horoball
  computation.
\item[\textbf{C7.}] \emph{Gluing assert suite.}
  Permutations, round-trips, class sizes, connectivity, volumes.  Check:
  \texttt{congruence\_prototype.py gluing} exits clean for all three
  levels.
\item[\textbf{C8.}] \emph{Self-ID relaxation.}
  $j=c$ faces are Neumann; $j\neq c$ faces are conforming.  Check:
  dof-graph connected components $=1$; $\ker Q\simeq\mathrm{const}$.
\item[\textbf{C9.}] \emph{Volume identity.}
  $(N\p+1)\vol(K)+\sum|T_\alpha|/(2Y^2)=(N\p+1)\vol(F)$.  Check:
  $1^TM1$ matches $(N\p+1)\vol_w(K)$ up to $O(h^2)$ sliver.
\end{enumerate}

\section{Gluing tables (summary)}

For each level, the right action of the four generators on
$\mathbb{P}^1(\ZZ[i]/\p)$ is a permutation of the $N\p+1$ cosets, with
cusp-class partition $\{1,N\p\}$.  The production code
(\texttt{GaussianResidueRing}) emits the full tables; we record only the
structural invariants used by the certificate:

\begin{center}
\begin{tabular}{@{}llll@{}}
\toprule
Level & $|\mathbb{P}^1|$ & Cusp $\infty$ & Cusp $0$ \\
\midrule
$(2+i)$ & $6$ & $1$ coset & $5$ cosets \\
$(3)$ & $10$ & $1$ coset & $9$ cosets \\
$(3+2i)$ & $14$ & $1$ coset & $13$ cosets \\
\bottomrule
\end{tabular}
\end{center}

Involutions: $S^2=\mathrm{id}$ and $R^2=T_iR^2=\mathrm{id}$ in $\PSL$.
Round-trip: acting by $\delta$ then $\delta^{-1}$ is the identity
permutation.  All three levels pass the assert suite.

\begin{thebibliography}{99}

\bibitem{BB11}
V.~Blomer and F.~Brumley,
\emph{On the Ramanujan conjecture over number fields},
Ann.\ of Math.\ (2) \textbf{174} (2011), 581--605.

\bibitem{EGM98}
J.~Elstrodt, F.~Grunewald, and J.~Mennicke,
\emph{Groups Acting on Hyperbolic Space},
Springer, 1998.

\bibitem{Friedman}
J.~Friedman,
\emph{The Selberg trace formula for $\mathrm{PSL}(2,O_K)$},
arXiv:math/0612807.

\bibitem{PaperI}
\emph{Exceptional eigenvalues of Bianchi orbifolds by verified finite
elements.  I: The method, and the spectral gap of the Picard orbifold},
companion paper (this series).

\bibitem{PaperIII}
\emph{Exceptional eigenvalues of Bianchi orbifolds by verified finite
elements.  III: The Eisenstein--Picard orbifold and its first congruence
cover},
companion paper (this series).

\bibitem{Rump06}
S.~M.~Rump,
\emph{Verification of positive definiteness},
BIT Numer.\ Math.\ \textbf{46} (2006), 433--452.

\end{thebibliography}

\end{document}
